Parsing Postfix log files

### 2009-03-05 doc/logparser-acronyms.tex
Removed URLs that are the first Google hit for a search term.

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language},
    first={Structured Query Language (SQL)}
}
\newglossaryentry{SQL glossary}{
    name={Structured Query Language},
    description={
        The standard language used for database querying, modification, and
        maintenance.  Some information and history can be found at
        \urlLastChecked{http://en.wikipedia.org/wiki/SQL}{2009/02/23}; a
        good introduction can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
    }
}

>>>>

\newglossaryentry{SQL}{
    type=\acronymtype,
    name={SQL},
    description={Structured Query Language},
    first={Structured Query Language (SQL)}
}

### 2009-03-05 doc/logparser-architecture.tex
Removed URLs that are the first Google hit for a search term.

Rules can specify conditions that will be evaluated by the framework before
attempting to use a rule to recognise an input: if the condition evaluates
to true the rule will be used, if not the rule will be skipped.  Conditions
can be as simple or complex as the parser requires, though naturally as the
complexity rises so does the difficulty in understanding how everything
interacts.  The framework has to evaluate each condition, so as the
complexity of conditions increases so does the complexity of the code
required to evaluate them; beyond a certain level of complexity conditions
should probably be written in a proper programming language, e.g.\ taking
advantage of dynamic languages' support for evaluating code at run-time, or
embedding a language like Lua
(\urlLastCheckedNoParens{http://ww.lua.org/}{2009/03/03}).  If an
implementation is going to use conditions so complex that they will require
a full programming language to implement them, the design may need to be
re-thought, including the choice to use this architecture --- there may be
other architectures more suitable.

@@

Usually excessive CPU time is consumed when a regex with a lot of
backtracking, caused by alteration and/or nested quantifiers, fails to
match; successful matching is generally quite fast with such regexes, so
problematic regexes may go unnoticed for some time.  E.g.\ with most regex
engines matching double quoted strings with \verb!"([^"\\]+|\\.)*"! is very
fast when there is a match, but when the match fails its computational
complexity is $O(2^{n})$; see~\cite{mastering-regular-expressions} for
in-depth discussion of nested quantifiers, backtracking, alteration, and
capturing groups.  Pathological regexes which consume excessive CPU time
can be difficult to detect, both by visual inspection and by machine
inspection, but if a regex is converted to a \acronym{FA} or the internal
representation used by the regex engine, it may be possible to determine if
there are nested quantifiers or other troublesome constructs present.
Modern regex engines have addressed many of these problems, e.g.\ the regex
to match double quoted strings given above fails immediately with Perl
5.10, regardless of the input length, because the regex engine looks for
the required double quotes first.  Similarly Perl 5.10's regex engine
optimises alterations starting with literal text into a trie, which has
constant matching time, whereas the matching time for alteration is
proportional to the number of alternatives.  Perl 5.10 regexes can use
\verb!(?>pattern)!, which matches \verb!pattern! the first time the regex
engine passes over it, and does not change what it originally matched if
the regex engine backtracks over it, so troublesome regexes can be modified
to use this to reduce the impact of backtracking; Prolog
(\urlLastCheckedNoParens{http://en.wikipedia.org/wiki/Prolog}{2009/03/03})
users will notice a similarity to the \verb'!' (cut) operator.  A
presentation showing some of Perl 5.10's regex features is available at
\urlLastChecked{http://www.regex-engineer.org/slides/perl510_regex.html}{2009/03/03}.

>>>>

Rules can specify conditions that will be evaluated by the framework before
attempting to use a rule to recognise an input: if the condition evaluates
to true the rule will be used, if not the rule will be skipped.  Conditions
can be as simple or complex as the parser requires, though naturally as the
complexity rises so does the difficulty in understanding how everything
interacts.  The framework has to evaluate each condition, so as the
complexity of conditions increases so does the complexity of the code
required to evaluate them; beyond a certain level of complexity conditions
should probably be written in a proper programming language, e.g.\ taking
advantage of dynamic languages' support for evaluating code at run-time, or
embedding a language like Lua.  If an implementation is going to use
conditions so complex that they will require a full programming language to
implement them, the design may need to be re-thought, including the choice
to use this architecture --- there may be other architectures more
suitable.

@@

Usually excessive CPU time is consumed when a regex with a lot of
backtracking, caused by alteration and/or nested quantifiers, fails to
match; successful matching is generally quite fast with such regexes, so
problematic regexes may go unnoticed for some time.  E.g.\ with most regex
engines matching double quoted strings with \verb!"([^"\\]+|\\.)*"! is very
fast when there is a match, but when the match fails its computational
complexity is $O(2^{n})$; see~\cite{mastering-regular-expressions} for
in-depth discussion of nested quantifiers, backtracking, alteration, and
capturing groups.  Pathological regexes which consume excessive CPU time
can be difficult to detect, both by visual inspection and by machine
inspection, but if a regex is converted to a \acronym{FA} or the internal
representation used by the regex engine, it may be possible to determine if
there are nested quantifiers or other troublesome constructs present.
Modern regex engines have addressed many of these problems, e.g.\ the regex
to match double quoted strings given above fails immediately with Perl
5.10, regardless of the input length, because the regex engine looks for
the required double quotes first.  Similarly Perl 5.10's regex engine
optimises alterations starting with literal text into a trie, which has
constant matching time, whereas the matching time for alteration is
proportional to the number of alternatives.  Perl 5.10 regexes can use
\verb!(?>pattern)!, which matches \verb!pattern! the first time the regex
engine passes over it, and does not change what it originally matched if
the regex engine backtracks over it, so troublesome regexes can be modified
to use this to reduce the impact of backtracking; Prolog users will notice
a similarity to the \verb'!' (cut) operator.  A presentation showing some
of Perl 5.10's regex features is available at
\urlLastChecked{http://www.regex-engineer.org/slides/perl510_regex.html}{2009/03/03}.

### 2009-03-05 doc/logparser-implementation.tex
Some minor formatting fixes. Documented some of 'make check'. Added check for sentences with only one space separating them.
Wrote the first half of the rule conditions section; it needs to be augmented with an example, and maybe some more explanation.
Removed URLs that are the first Google hit for a search term.

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \acronym{SQL} can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23},
        others are
        \urlLastChecked{http://www.w3schools.com/sql/default.asp}{2009/02/23},
        \urlLastChecked{http://sqlcourse.com/}{2009/02/23}).  Although
        every vendor implements a different dialect of \acronym{SQL}, the
        basics are the same everywhere (analogous to the overall
        similarities and minor differences amongst Irish English, British
        English, American English, and Australian English).  Depending on
        the database in use there may be tools available that reduce or
        remove the requirement to know \acronym{SQL}; several are available
        for \gls{SQLite}, the default database used by the parser
        implementation.

@@

Each log line contains the name of the Postfix component that produced it,
and this information is used when determining which rules should be used to
parse each log line (see \sectionref{rules in implementation} for details).

>>>>

    \item \acronym{SQL} is reasonably standard and many people will already
        be familiar with it; for those unfamiliar with it there are lots of
        readily available resources from which to learn (a good
        introduction to \acronym{SQL} can be found at
        \urlLastChecked{http://philip.greenspun.com/sql/}{2009/02/23}.
        Although every vendor implements a different dialect of
        \acronym{SQL}, the basics are the same everywhere (analogous to the
        overall similarities and minor differences amongst Irish English,
        British English, American English, and Australian English).
        Depending on the database in use there may be tools available that
        reduce or remove the requirement to know \acronym{SQL}; several are
        available for \gls{SQLite}, the default database used by the parser
        implementation.

@@

XXX ONCE THIS PARAGRAPH HAS BEEN INTEGRATED AND EXPANDED UPDATE THE
REFERENCE IN ARCHITECTURE->RULES->ATTACHING CONDITIONS TO RULES\@.
Each log line contains the name of the Postfix component that produced it,
and this information is used when determining which rules should be used to
parse each log line (see \sectionref{rules in implementation} for details).

### 2009-03-05 doc/logparser-introduction.tex
Removed URLs that are the first Google hit for a search term.

The architecture described herein was developed as part of a larger project
to improve anti-spam defences by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are content-based
(e.g.~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}) and
require the mail to be accepted before determining if it is spam, but
rejecting mail during the delivery attempt is preferable: senders of
non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; resource usage is reduced on the accepting mail server
(allowing more intensive content-based techniques to be used on the mail
that is accepted); users have less spam mail to wade through.  Improving
the performance of anti-spam techniques that are applied when mail is being
transferred via \acronym{SMTP} is the goal of this project, by providing a
platform for reasoning about anti-spam filters.  The approach chosen to
measure performance is to analyse the log files produced by the
\acronym{SMTP} server in use, Postfix
(\urlLastCheckedNoParens{http://www.postfix.org/}{2009/02/28}), rather than
modifying its source code to generate statistics: this approach improves
the chances of other Postfix users testing and using the software developed
for this project.  The need arose for a parser capable of dealing with the
great number and variety of log lines produced by Postfix: the parser must
be designed so that adding support for parsing new inputs is a simple task,
because the log lines to be parsed will change over time.

>>>>

The architecture described herein was developed as part of a larger project
to improve anti-spam defences by analysing the performance of the set of
filters currently in use, optimising the order and membership of the set
based on that analysis, and developing supplemental filters where
deficiencies are identified.  Most anti-spam techniques are content-based
(e.g.~\cite{a-plan-for-spam, word-stemming, relaxed-online-svms}) and
require the mail to be accepted before determining if it is spam, but
rejecting mail during the delivery attempt is preferable: senders of
non-spam mail that is mistakenly rejected will receive an immediate
non-delivery notice; resource usage is reduced on the accepting mail server
(allowing more intensive content-based techniques to be used on the mail
that is accepted); users have less spam mail to wade through.  Improving
the performance of anti-spam techniques that are applied when mail is being
transferred via \acronym{SMTP} is the goal of this project, by providing a
platform for reasoning about anti-spam filters.  The approach chosen to
measure performance is to analyse the log files produced by the
\acronym{SMTP} server in use, Postfix, rather than modifying its source
code to generate statistics: this approach improves the chances of other
Postfix users testing and using the software developed for this project.
The need arose for a parser capable of dealing with the great number and
variety of log lines produced by Postfix: the parser must be designed so
that adding support for parsing new inputs is a simple task, because the
log lines to be parsed will change over time.

### 2009-03-05 doc/logparser-postfix-daemons.tex
Removed URLs that are the first Google hit for a search term.

\newglossaryentry{sendmail}{name={sendmail},type={postfix},description={
    A Postfix component that is compatible with the Sendmail
    (\urlLastCheckedNoParens{http://www.sendmail.org/}{2009/02/23}) mail
    submission program which all Unix commands that need to send mail use;
    it executes \daemon{postdrop} to place a new mail in the queue.
    \daemonDocURL[\newline{}]{http://www.postfix.org/sendmail.1.html}{2009/02/23}.
}}

>>>>

\newglossaryentry{sendmail}{name={sendmail},type={postfix},description={
    A Postfix component that is compatible with the Sendmail mail
    submission program which all Unix commands that need to send mail use;
    it executes \daemon{postdrop} to place a new mail in the queue.
    \daemonDocURL[\newline{}]{http://www.postfix.org/sendmail.1.html}{2009/02/23}.
}}
